---
title: "HEK293T_Mengrui"
author: "MVI"
date: "2024-02-02"
output: html_document
---

Load in packages
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("DESeq2")
BiocManager::install("arrayQualityMetrics")
install.packages("gplots")
install.packages("pheatmap")
install.packages("vegan")
install.packages("ggplot2")
install.packages("ggrepel")
install.packages("VennDiagram")
install.packages("ggtext")
install.packages("ggpubr")
install.packages("reshape2")
install.packages("tidyselect")
install.packages("dplyr")

library(DESeq2)
library(arrayQualityMetrics)
library(gplots)
library(pheatmap)
library(vegan)
library(ggplot2)
library(ggrepel)
library(VennDiagram)
library(ggtext)
library(ggpubr)
library(reshape2)
library(tidyselect)
library(dplyr)
```

Set your working directory (the entire path to the folder that contains this rmarkdown and your data files)
```{r}
knitr::opts_knit$set(root.dir = "/Users/mariaingersoll/Desktop/BU_Research/Davies-Gilmore/Mengrui_HEK293T")
#using the knitr option, it ensures that all chunks run in this rmd are performed in this working directory

#check your working directory:
getwd()
```

Open and inspect counts data (preprocessed using "tagseq_processing_HOWTO" on BU SCC)
```{r}
#load the count file into R and name it "countData"
countData = read.table("./Mengrui_HEK293T_2024_counts.txt")

#how long is the file (in other words, how many genes do you have count data for)?
length(countData[,1])
#59048

#look at the first few rows of the your file
head(countData)

#rename each of your samples (each column) with shorter names
newColNames = c("flag1_S33", "flag2_S34", "flag3_S35", "N1_S36", "N2_S37", "N3_S38", "NLS1_S39", "NLS2_S40", "NLS3_S41")
colnames(countData)=paste(newColNames)
length(countData[,1])
#59048
head(countData)

#look at the distribution of your count data throughout your samples
totalCounts=colSums(countData)
barplot(totalCounts, col=c("mediumblue", "mediumblue", "mediumblue", "red", "red", "red", "darkgreen", "darkgreen", "darkgreen"), ylab="Raw Counts", xlab="HEK293T")


min(totalCounts)
#480525     
max(totalCounts)
#862648 
```

Need to make a metadata file with information about each sample and its respective condition
```{r}
sample=c("flag1_S33", "flag2_S34", "flag3_S35", "N1_S36", "N2_S37", "N3_S38", "NLS1_S39", "NLS2_S40", "NLS3_S41")
s=data.frame(sample)
sampledata <-s
sampledata

treat=c("Flag", "Flag", "Flag", "N", "N", "N", "NLS", "NLS", "NLS")
g=data.frame(treat)
colData<- g
colData

allData = cbind(sampledata, colData)
allData
```

Run preliminary DESeq (differential expression testing) and outlier analysis (using arrayqualitymetrics)
```{r}
dds<-DESeqDataSetFromMatrix(countData=countData, colData=colData, design=~treat)
dds<-DESeq(dds)

vsd.ge=assay(vst(dds))
rl=vst(dds)
e=ExpressionSet(assay(rl), AnnotatedDataFrame(as.data.frame(colData(rl))))
v = setwd("~/Desktop/BU_Research/Davies-Gilmore/Mengrui_HEK293T")
arrayQualityMetrics(e,outdir=v,intgroup=c("treat"),force=T)
```
Open the index.html file that is printed into the output directory
--> no outliers

Visualize the results
`results` extracts the results table from the DESeq analysis and the DESeqDataSet is used to generate the dispersion plot.
  --> Have to compare each of your "treatments" to the control
  --> And then the "treatments" to each other too
  
```{r}
#N to Flag
head(dds)
res_N_Flag<- results(dds, contrast = c("treat", "N", "Flag"))
res_N_Flag

#NLS to Flag
res_NLS_Flag<- results(dds, contrast = c("treat", "NLS", "Flag"))
res_NLS_Flag

#N to NLS
res_N_NLS<- results(dds, contrast = c("treat", "N", "NLS"))
res_N_NLS
```

The above code gives the log2 fold change (MLE) of a treatment against a baseline
For example, in res_N_Flag, a pos log2FC values indicates that a gene is upregulated in N compared to Flag
-->The name provided in the second element of the results() command is the baseline


Dispersion plot
Read more about dispersion here: https://hbctraining.github.io/DGE_workshop/lessons/04_DGE_DESeq2_analysis.html

Basically, as the mean of normalized counts for a given gene increases, the dispersion (variance) should decrease
```{r}
plotDispEsts(dds, main="Dispersion Uptake")
```

- Now we are going to retrieve the rlog data, which will give us better transformation when size factors vary across samples
- regularized log transformation, transforms COUNT DATA to log2 scale to minimize differences between samples for rows with small counts
```{r}
rld <- rlogTransformation(dds, blind=TRUE)
head(assay(rld))
hist(assay(rld))
```

- The assay function allows you to access the matrix-like data, so that you can view it bc head(rld) gives you just a summary
- The histogram above displays the frequency of each binned rld count value

Raw Sample Comparison

- Making a sample distance heatmap
- as.matrix attempts to turn its argument into a matrix
- dist computes and returns the distance matrix computed using the specified distance measure to compute distances between the rows of a data matrix (assay(rld))
- Create a sample heatmap of the relatedness based on the count distance/difference between samples
```{r}
sampleDists <- as.matrix(dist(t(assay(rld))))

heatmap.2(as.matrix(sampleDists), key=F, trace="none",
          col=colorpanel(100, "black", "white"), ColSideColors=c("mediumblue",  "mediumblue", "mediumblue", "red", "red", "red", "darkgreen", "darkgreen", "darkgreen"),RowSideColors = c("mediumblue", "mediumblue", "mediumblue", "red", "red", "red", "darkgreen", "darkgreen", "darkgreen"),
          margin=c(10, 10))

#saved as sampledisthm.pdf as landscape with 5x5 dimensions

#all the flags are grouping together and then the Ns and the NLS are grouping together
```


